### 2017-01-27 manuscript.rmd
Changes wording

The first study, Kryuchkova-Mostacci and Robinson-Rechavi (KMRR) 2016 [@KryuchkovaMostacci:2016iw], analyzed multiple vertebrate expression datasets to test the ortholog conjecture - the hypothesis that orthologs tend to have more conserved attributes (specificity of expression across organs in this case) than do paralogs [@Nehrt:2011bm]. Using pairwise comparisons (Fig. 1A), they found lower expression correlation between paralogs than between orthologs and interpreted this as strong support for the ortholog conjecture.

@@

	# These clade dates are from the KMRR code, they got them from http://timetree.org/
	calibration_times = data.frame(
		age.min =
			c(  8, 29, 92, 162, 167, 296 ), 
		clade = 
			c( "Homininae", "Catarrhini", "Euarchontoglires", "Theria", "Mammalia", "Amniota" ),
		stringsAsFactors=FALSE
	)
```

@@

	#' Adds some random noise from a normal distribution to calibration
	#' times
	#' 
	#' @param calibration_times A dataframe with two columns: age.min is the age 
	#' of a clade, and clade is the name of the clade
	#' @param sd_fraction this is multiplied by each value to set sd of that value 
	#' @return A dataframe as above, but with noise
	#' @export
	add_noise_calibration_times = function( calibration_times, sd_fraction ) {
		age = calibration_times$age.min
		new_age = NA

		while( NA %in% new_age ) {
			x = sapply( 
				age,
				function( i ){
					rnorm( 1, mean = i, sd = ( sd_fraction * i ) )
				}
			)

			if( ! is.unsorted( x ) ){
				new_age = x
			}

		calibration_times$age.min = new_age
		return( calibration_times )
	}

@@

		tree = try( 
			ape::chronos( nhx@phylo, calibration=calibration, ... ) 
		)

@@

	#' Calibrate a list of gene trees given a set of speciation node dates
	#' 
	#' @param gene_trees_pruned A list of phylogenetic trees as treeio::treedata objects 
	#' with associated character data
	#' @param calibration_times A dataframe with two columns: age.min is the age 
	#' of a clade, and clade is the name of the clade
	#' @param ... Any additional arguments to pass to ape::chronos()
	#' @return A list of calibrated phylogenetic trees as treeio::treedata objects. May be 
	#' shorter than gene_trees_pruned if some calibrations fail.
	#' @export
	calibrate_trees = function( gene_trees_pruned, calibration_times, ... ) {
		# Make the trees ultrametric and calibrate the speciation nodes to 
		# specified times
		gene_trees_calibrated = 
			mclapply( 
				gene_trees_pruned,
				calibrate_tree, 
				calibration_times=calibration_times,
				...,
				mc.cores=cores
			)

@@

	# Remove tips from the trees that do not have expression data
	gene_trees_pruned = mclapply( 
		gene_trees_annotated, 
		drop_empty_tips, 
		min_genes_with_expression=min_genes_with_expression,
		mc.cores=cores 
	)

	# Remove trees with less than min_genes_with_expression tips with expression data
	gene_trees_pruned = gene_trees_pruned[ ! is.na( gene_trees_pruned ) ]

@@

	# Time calibrate all speciation nodes to the same years
	gene_trees_calibrated = calibrate_trees( gene_trees_pruned, calibration_times, model="correlated" )

@@

When we reanalyze the data with phylogenetic independent contrasts (Fig. 1B), we did not find increased evolutionary change in expression following duplication events (Fig. XXKMRR). Of the `r length(gene_trees)` trees that were parsed from ENSMBL Compara [@Herrero:2016jj], `r length(gene_trees_pruned)` passed taxon sampling criteria (`r min_genes_with_expression` genes) after removing tips without Tau values. Of these, `r length(gene_trees_calibrated)` were successfully time calibrated. These calibrated pruned trees were used to calculate phylogenetic independent contrasts for Tau. They contain `r nodes_contrast %>% filter(D=="Y") %>% nrow()` duplication nodes and `r nodes_contrast %>% filter(D=="N") %>% nrow()` speciation nodes.

@@

Recent advances in phylogenetic comparative methods are particularly well suited to addressing questions about the evolution of functional genomic traits. Most early phylogenetic comparative methods attempted to remove evolutionary signal to correct statistical tests for correlations between traits, while more recent methods tend to focus on testing hypotheses of evolutionary processes [@Pennell:2013cw]. The application of this newer focus to functional genomics provides an exciting opportunity to address long standing questions of broad interest, including the order of changes in functional genomics traits and shifts in rates of evolution of one functional genomics trait following changes in another trait.

>>>>

The first study, Kryuchkova-Mostacci and Robinson-Rechavi (KMRR) 2016 [@KryuchkovaMostacci:2016iw], analyzed multiple vertebrate expression datasets to test the ortholog conjecture - the hypothesis that orthologs tend of have more conserved attributes (specificity of expression across organs in this case) than do paralogs [@Nehrt:2011bm]. Using pairwise comparisons (Fig. 1A), they found lower expression correlation between paralogs than between orthologs and interpreted this as strong support for the ortholog conjecture.

@@

	# These clade dates are from the KMRR code, they got them from http://timetree.org/
	calibration_times = data.frame(
		age.min =
			c( 92, 167, 8, 296, 29, 162 ), 
		clade = 
			c( "Euarchontoglires", "Mammalia", "Homininae", 
					"Amniota", "Catarrhini", "Theria"
			),
		stringsAsFactors=FALSE
	)
```

@@

@@

		tree = try( 
			ape::chronos( nhx@phylo, calibration=calibration ) 
		)

@@

	#' Calibrate a list of gene trees given a set of speciation node dates
	#' 
	#' @param gene_trees_sub A list of phylogenetic trees as treeio::treedata objects 
	#' with associated character data
	#' @param calibration_times A dataframe with two columns: age.min is the age 
	#' of a clade, and clade is the name of the clade
	#' @param ... Any additional arguments to pass to ape::chronos()
	#' @return A list of calibrated phylogenetic trees as treeio::treedata objects. May be 
	#' shorter than gene_trees_sub if some calibrations fail.
	#' @export
	calibrate_trees = function( gene_trees_sub, calibration_times, ... ) {
		# Make the trees ultrametric and calibrate the speciation nodes to 
		# specified times
		gene_trees_calibrated = 
			mclapply( 
				gene_trees_sub,
				calibrate_tree, 
				calibration_times=calibration_times,
				...,
				mc.cores=cores
			)

@@

	# Remove tips from the trees that do not have expression data
	gene_trees_sub = mclapply( 
		gene_trees_annotated, 
		drop_empty_tips, 
		min_genes_with_expression=min_genes_with_expression,
		mc.cores=cores 
	)

	# Remove trees with less than min_genes_with_expression tips with expression data
	gene_trees_sub = gene_trees_sub[ ! is.na( gene_trees_sub ) ]

@@

	# Time calibrate all speciation nodes to the same years
	gene_trees_calibrated = calibrate_trees( gene_trees_sub, calibration_times, model="correlated" )

@@

When we reanalyze the data with phylogenetic independent contrasts (Fig. 1B), we did not find increased evolutionary change in expression following duplication events (Fig. XXKMRR). Of the `r length(gene_trees)` trees that were parsed from ENSMBL Compara [@Herrero:2016jj], `r length(gene_trees_sub)` passed taxon sampling criteria (`r min_genes_with_expression` genes) after removing tips without Tau values. Of these, `r length(gene_trees_calibrated)` were successfully time calibrated. These calibrated pruned trees were used to calculate phylogenetic independent contrasts for Tau. They contain `r nodes_contrast %>% filter(D=="Y") %>% nrow()` duplication nodes and `r nodes_contrast %>% filter(D=="N") %>% nrow()` speciation nodes.

@@

Recent advances in phylogenetic comparative methods are particularly well suited to addressing questions about the evolution of functional genomic traits. Most early phylogenetic comparative methods attempted to account for evolutionary signal to correct statistical tests for correlations between traits, while more recent methods tend to focus on testing hypotheses of evolutionary processes [@Pennell:2013cw]. The application of this newer focus to functional genomics provides an exciting opportunity to address long standing questions of broad interest, including the order of changes in functional genomics traits and shifts in rates of evolution of one functional genomics trait following changes in another trait.
